Manuscript Assignment = Calcasieu

### 2025-04-18 Manuscript_Assignment.qmd
c

library(tidyverse)
library(hypervolume)
library(MuMIn)

@@

@@

@@

@@

>>>>

library(tidyverse)
library(hypervolume)
library(MuMIn)
library(ggpubr)
library(gt)

@@

## HV plots

@@

## Niche SIze

Nice size of the three species

```{r species - plot size}
#| echo: FALSE

dfs_hvs |>
  select(species, size) |>
  ggplot(aes(x = species, y = 1, size = size, fill = species)) +
  geom_point(shape = 21, color = "black") +
  geom_text(aes(label = round(size, 2)), vjust = -3, size = 9) +  # nicely formatted labels
  scale_size_continuous(range = c(10, 40)) +
  labs(
    title = "Niche Size of Estuarine Species",
    x = "Species",
    y = NULL
  ) +
  guides(size = "none", fill = "none") +  # remove legends
  theme_classic(base_size = 14) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

```{r species - table overlap}
#| echo: FALSE
#| eval: FALSE

df_ov_dedup <- df_ov |> 
  group_by(y1, y2) |> 
  summarise(across(where(is.numeric), mean), .groups = "drop")

df_ov_dedup |>
  mutate(across(where(is.numeric), ~ round(., 3))) |>
  gt() |>
  tab_header(
    title = "Pairwise Niche Overlap and Dissimilarity Metrics"
  ) |>
  cols_label(
    y1 = "Species 1",
    y2 = "Species 2",
    lsr = "Log Size Ratio",
    dist_cent = "Centroid Distance",
    jaccard = "Jaccard",
    sorensen = "Sorensen",
    uniq_y1 = "Unique to Species 1",
    uniq_y2 = "Unique to Species 2"
  )

```
![Pairwise Niche Overlap Table](images/niche_overlap_table.png)

Heatmap of similarity between species using Sorensen

![Niche Overlap Heatmap](sorensen_heatmap.png)

@@

## Habitat Suitability/Time

@@

## Species Abundance

I went ahead and plotted abundance data. No brown shrimp or white shimp were caught in 2019 and in general there was a continous decrease in their abundance numbers.

We can see that sharp fluctuation in abundance is normal but did not reach zero before 2019. Panel B just has a closer look at the last three years

```{r - abundance}
#| echo: TRUE

target_years <- as.character(1986:2019)
target_species <- c("white shrimp", "brown shrimp", "blue crab")

# Fill in missing combinations with 0
species_counts <- df |>
  filter(species %in% target_species,
         year %in% target_years) |>
  group_by(year, species) |>
  summarise(n_caught = sum(num), .groups = "drop") |>
  complete(year = target_years, species = target_species, fill = list(n_caught = 0)) |>
  arrange(year, species)

a = ggplot(species_counts, aes(x = as.numeric(year), y = n_caught, color = species)) +
  geom_line(size = 1.2) +
  geom_point(size = 2.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Abundance of Estuarine Species (1986–2019)",
    x = "Year",
    y = "Number Caught",
    color = "Species"
  ) +
  theme_classic(base_size = 14)

species_counts_zoom <- species_counts |>
  filter(as.numeric(year) >= 2017 & as.numeric(year) <= 2019)

b = ggplot(species_counts_zoom, aes(x = as.numeric(year), y = n_caught, color = species)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = 2017:2019, labels = 2017:2019) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Abundance of Estuarine Species (2017–2019)",
    x = "Year",
    y = "Number Caught",
    color = "Species"
  ) +
  theme_classic(base_size = 14)

ggarrange(a, b,
          ncol = 2, nrow = 1,
          labels = c("A", "B"),
          common.legend = TRUE,
          legend = "bottom") +
  theme(plot.title = element_text(hjust = 0.5))

## 2019 Suitability

Upon Further investigation. we see that the habitat suitbability in 2019 changed outsize of the centroid of white shrimp and brown shrimp. It was not longer suitable and had greater changes in magnitude compared to blue crab.....(i think)

```{r - suitability - 2019}
#| echo: FALSE
#| message: false

suitability_2019_summary = read_csv("data/suitability_2019_summary.csv")

suitability_2019_summary |> 
  mutate(across(where(is.numeric), ~ round(., 2))) |>
  gt() |>
  tab_header(
    title = "Habitat Suitability and Niche Deviation (2019)"
  ) |>
  cols_label(
    species = "Species",
    prop_suitable = "Prop. Suitable",
    salinity = "Δ Salinity",
    watertemp = "Δ Water Temp",
    turbidity = "Δ Turbidity",
    airtemp = "Δ Air Temp"
  ) |>
  fmt_number(everything(), decimals = 2)
